Introductory statistics with `intRo`

### 2016-06-15 jcgs-paper-2016/paper.Rmd
more comments

1. Course instructors can easily enable or disable functionality depending on the exact needs of the particular class being taught. Because this is done at run-time, including new modules is as simple as downloading and placing it within 'intRo's` modules folder, or removing existing modules from that folder.
2. Errors can be more easily isolated to specific components. For instance, if an error is encountered, simply disabling the module can provide a temporary workaround while the issue is identified.
3. Modularity helps to organize the different pieces of code into functionality chunks that make it easier for developers to maintain.

@@

The modules provided with `intRo` are contained in the modules folder. The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `statistics`). Within each of these categories is a folder named according to the name of the module. This folder houses the previously defined scripts. As an example, we will walk through the contents of the `contingency` module.

Since the `contingency` module performs a statistical analysis, it is part of the statistics category, and hence can be found in the `intRo` repository at **modules/statistics/contingency**. Let's first examine *helper.R*:

@@

Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by Shiny, results automatically update when users interact with the interface.

@@

@@

We believe `intRo` can be a powerful and effective tool for introductory statistics education. Its modular structure allows it to be flexible enough for many different applications and curriculums. Its ease-of-use allows the student to focus her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation. Reproducible code generated from each analysis can be used to spark an interest in R programming in those who might otherwise not be exposed to it.

@@

Challenges do exist with regards to the wider adoption of `intRo`. For instance, we will need to monitor how well the server hosting `intRo` handles the load of dozens of students performing data analyses at once. If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load. An unknown quantity will be how feasible it is to increase adoption of `intRo` across the department, as well as to other universities. One limitation of `intRo` is that uploading a dataset beyond about 30,000 rows tends to be slow. Even once the data is successfully uploaded, the default modules produce results more slowly than with smaller datasets. This is a limitation that should be further investigated if and when `intRo` sees wider adoption.

>>>>

1. Course instructors can easily enable or disable functionality depending on the exact needs of the particular class being taught. Because this is done at run-time, including new modules is as simple as downloading and placing it within `intRo's` modules folder, or removing existing modules from that folder.
2. Errors can be more easily isolated to specific components. For instance, if an error is encountered, simply disabling the module can provide a temporary workaround while the issue is identified.
3. Modularity helps to organize the different pieces of code into functionality chunks that make it easier for developers to maintain.

@@

The modules provided with `intRo` are contained in the modules folder. The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `inference`). Within each of these categories is a folder named according to the name of the module. This folder houses the previously defined scripts. As an example, we will walk through the contents of the `contingency` module.

**AK: Should we walk through a difference models? Since this is not inference, it is confusing**
Since the `contingency` module performs a statistical analysis, it is part of the inference category, and hence can be found in the `intRo` repository at **modules/inference/contingency**. Let's first examine *helper.R*:

@@

Reactive programming is a programming paradigm that "tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies" [@bainomugisha2012survey]. As implemented by Shiny, results automatically update when users interact with the interface.

@@

**AK: I feel like this section could use a little more beef, but I'm not sure how**

@@

We believe **AK: we believe is weak language, consider removing** `intRo` can be a powerful and effective tool for introductory statistics education. Its modular structure allows it to be flexible enough for many different applications and curriculums. Its ease-of-use allows the student to focus her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation. Reproducible code generated from each analysis can be used to spark an interest in R programming in those who might otherwise not be exposed to it.

@@

Challenges do exist with regards to the wider adoption of `intRo`. For instance, we will need to monitor how well the server hosting `intRo` handles the load of dozens of students performing data analyses at once. If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load. An unknown quantity will be how feasible it is to increase adoption of `intRo` across the department **AK: not sure about "the department"**, as well as to other universities. One limitation of `intRo` is that uploading a dataset beyond about 30,000 rows tends to be slow. Even once the data is successfully uploaded, the default modules produce results more slowly than with smaller datasets. This is a limitation that should be further investigated if and when `intRo` sees wider adoption.
